Query Optimization for Dynamic Imputation

### 2017-05-12 rebuttal.tex

  \item D8. Are there impute implementation that are non-blocking? Or do they always have to perform multiple passes over the data?

    \resp{Yes. For example, in time series data, it is reasonable to fill the most recent non-null value forward. This kind of imputation isn't useful for survey data, so we didn't consider it for our experiments. We have extended the experiments section to consider a simple non-block imputation method relevant to our dataset: mean value imputation. This is a common technique, which fills in  missing values
    with the average observed value. In our case, we implement a non-blocking variant that estimates the mean from the plan's histograms, avoiding the need to perform a full pass over
    the data before imputing.}

  \item D9. Additional experiments I suggest to include:
    -- Consider a scenario where the user submits multiple similar queries. Show after how many such queries your approach has no longer a competitive advantage compared to performing imputation on the data.
    -- Try different values for $\alpha$ (not just 0 and 1).
    -- Show runtime of enumeration algorithm as queries get more complex.

    \resp{
      The sensitivity of different values of $\alpha$ depends on the query itself. Performing a scan of values of $\alpha$ with small step sizes can return
      a lot of the same plans for execution. In lieu of this, we extended the results section to include a subsection that presents the final frontiers for
      two queries, one from each dataset used during evaluation. This frontier can be exposed to the user to help in their workflow.
      \\
      We explored the planning times for queries with increasing number of joins. The standard algorithm presented in the original version submitted handles
      up to 5 joins, with columns requiring imputation in all tables, with planning time of approximately one second or less. We added a small optimization to help reduce
      planning times for a larger number of joins. We maintain an approximation of the Pareto sets, rather than the exact set, by only adding non-dominated plans that
      are sufficiently different. This optimization reduces planning time linearly with respect to the reduction in the size of the final frontier. Users can increase
      the level of approximation to reduce planning times further.}

### 2017-05-12 paper/experiments.tex
added planning time plot etc to paper and discussion in rebuttal

\subsubsection{Planning Times}
\begin{table}
\centering
\input{tables/planning_summary_exact_pareto.tex}
\caption{Planning times are sub-second for up to 4 joins. For 5+  joins planning times rise. Many practical queries
can be formulated with a smaller number of joins, but to support more complex queries \ProjectName{} provides an
optimization that maintains approximate Pareto sets for queries with many joins. This reduces the 29 second
planning time shown for 7 joins by approximately 45\%, on average.
\todo{update with new table and values}
}
\label{table:planning-times}
\end{table}

\todo{Update: Planning times are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.}
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated increasing number of joins across tables.
Planning times for queries involving 1--5 joins, which is a practical value for real world exploratory queries, take less than a second.
\todo{update: For queries with 7 joins, planning time increases substantially to 29 seconds, on average. We introduced an optimization that reduces
the plan space by maintaining only approximations of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).}

%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}

>>>>

\subsubsection{Planning Times}
\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/planning_times_plot.png}
\caption{Planning times are sub-second for up to 5 joins. For more joins, we implement an approximation to Pareto sets.
This reduces planning time linearly with respect to the reduction in the final frontier size. Users can increase the level of
approximation for more aggressive planning reductions.
}
\label{fig:planning-times}
\end{figure}

\todo{Update: Planning times for our benchmark queries are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.}
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
\Cref{fig:planning-times} shows the average planning time on a log scale for queries of the CDC data with an increasing number of joins. As the number of tables requiring imputation grows,
so does the search space and corresponding planning time. Planning time for queries involving 1--5 joins are sub-second. While a useful subset of exploratory queries can be expressed
with a smaller number of joins, we extended \ProjectName{} to support an approximation of Pareto sets to improve planning times for queries with 6 or more joins. Approximate Pareto sets
displayed a reduction in planning time linear with respect to the reduction in size of the final frontier. Users can increase the level of approximation to reduce planning times further.
%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}
